Text embedding models yield high-resolution insights into conceptual knowledge from short multiple-choice quizzes

### 2025-01-09 paper/main.tex
minor typo fixes

In the set of analyses reported in Figure~\ref{fig:predictions}, we assessed
whether estimates of participants' knowledge at the embedding coordinates of
individual quiz questions could be used to reliably predict their ability to
correctly answer those questions. In essence, we treated each question a given
participant answered on a given quiz as a ``lecture'' consisting of a single
timepoint, and used Equation~\ref{eqn:prop} to estimate the participant's
knowledge for its embedding coordinate based on their performance on all
\textit{other} questions they answered on that same quiz (``All questions'';
Fig.~\ref{fig:predictions}, top row). Additionally, for each lecture-related
question (i.e., excluding questions about general physics knowledge), we
computed analogous knowledge estimates based on two different subsets of
questions the participant answered on the same quiz: (1) all \textit{other}
questions about the same lecture as the target question (``Within-lecture'';
Fig.~\ref{fig:predictions}, middle rows), and (2) all questions about the other
of the two lectures (``Across-lecture''; Fig.~\ref{fig:predictions},
bottom rows).

@@

In our application, all of the coordinates we embedded were topic vectors,
whose elements are always non-negative and sum to one. Although UMAP is an
invertible transformation at the embedding locations of the original data,
other locations in the embedding space will not necessarily follow the same
implicit ``rules'' as the original high-dimensional data. For example,
inverting an arbitrary coordinate in the embedding space might result in
negative-valued vectors, which are incompatible with the topic modeling
framework. To protect against this issue, we log-transformed the topic vectors
prior to embedding them in the 2D space. When we inverted the embedded vectors
(e.g., to estimate topic vectors or word clouds, as in
Fig.~\ref{fig:knowledge-maps}C), we passed the inverted (log-transformed)
values through the exponential function to obtain a vector of non-negative
values, and normalized them to sum to one.

>>>>

In the set of analyses reported in Figure~\ref{fig:predictions}, we assessed
whether estimates of participants' knowledge at the embedding coordinates of
individual quiz questions could be used to reliably predict their abilities to
correctly answer those questions. In essence, we treated each question a given
participant answered on a given quiz as a ``lecture'' consisting of a single
timepoint, and used Equation~\ref{eqn:prop} to estimate the participant's
knowledge for its embedding coordinate based on their performance on all
\textit{other} questions they answered on that same quiz (``All questions'';
Fig.~\ref{fig:predictions}, top row). Additionally, for each lecture-related
question (i.e., excluding questions about general physics knowledge), we
computed analogous knowledge estimates based on two different subsets of
questions the participant answered on the same quiz: (1) all \textit{other}
questions about the same lecture as the target question (``Within-lecture'';
Fig.~\ref{fig:predictions}, middle rows), and (2) all questions about the other
of the two lectures (``Across-lecture''; Fig.~\ref{fig:predictions},
bottom rows).

@@

In our application, all of the coordinates we embedded were topic vectors,
whose elements are always non-negative and sum to one. Although UMAP is an
invertible transformation at the embedding locations of the original data,
other locations in the embedding space will not necessarily follow the same
implicit ``rules'' as the original high-dimensional data. For example,
inverting an arbitrary coordinate in the embedding space might result in
negative-valued vectors, which are incompatible with the topic modeling
framework. To protect against this issue, we log-transformed the topic vectors
prior to embedding them in the 2D space. When we inverted the embedded vectors
(e.g., to estimate topic vectors for word clouds, as in
Fig.~\ref{fig:knowledge-maps}C), we passed the inverted (log-transformed)
values through the exponential function to obtain a vector of non-negative
values, and normalized them to sum to one.
